PANCHANGE_data_preprocessing

### 2020-08-11 PANCHANGE_data_preprocessing2020-07-15.Rmd
if statement added for phq prepan

```{r glad pcl copy}
if(GLAD == TRUE) {
  prepan.pcl.raw.id <- glad.pcl.raw.id
}
```

@@

```{r edgi pcl copy}
if(EDGI == TRUE) {
  prepan.pcl.raw.id <- edgi.pcl.raw.id
}
```

@@

```{r glad phq cols}
glad.phq.raw <- readRDS(file = paste0(data.raw_path, "/glad/phq_glad.rds"))
dim(glad.phq.raw)
colnames(glad.phq.raw)
```

@@

@@

*Note that scores with missingness will be changed to NA later in the script
```{r phq sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
phq.items_key_prepan <- c(1,1,1,1,1,1,1,1,1)

@@

# Calulate sum score
phq.scored_items_prepan <- scoreItems(keys = phq.items_key_prepan,
                                        items = glad.phq.raw.id[phq.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 3)
# Add in column to tibble
glad.phq.raw.id$phq.sum_score_prepan_raw <- phq.scored_items_prepan$scores

# Look at values
glad.phq.raw.id %>%
  descr(phq.sum_score_prepan_raw,
        stats = "common"
        )
```

@@

prepandemic
```{r phq prepan missingness}
glad.phq.raw.id <- glad.phq.raw.id %>%
  mutate(
    na_per_person_phq_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[phq.items_prepan])
      )
  )

glad.phq.raw.id %>%
  freq(na_per_person_phq_prepan,
       cumul = F)
```

```{r phq prepan missingness 8 items}
glad.phq.raw.id <- glad.phq.raw.id %>%
  mutate(
    phq.missing_only_suicide_item_prepan = 
      if_else(
        condition = na_per_person_phq_prepan == 1 & 
          is.na(phq.dead_hurting_thoughts_prepan_numeric), 
        true = 1,
        false = 0,
        missing = NA_real_
      )
  )

glad.phq.raw.id %>%
  freq(phq.missing_only_suicide_item_prepan,
       cumul = F)
```

all times
```{r phq prepan sum score if missigng items set to NA - all items}
glad.phq.raw.id <- glad.phq.raw.id %>%
  mutate(
    phq.sum_score_prepan =
      if_else(condition = na_per_person_phq_prepan > 0,
              true = NA_real_,
              false = phq.sum_score_prepan_raw,
              missing = NA_real_
      )
  )

glad.phq.raw.id %>%
  descr(phq.sum_score_prepan_raw, stats = "common")

glad.phq.raw.id %>%
  descr(phq.sum_score_prepan, stats = "common")
```

>>>>

```{r glad pcl prepan copy}
if(GLAD == TRUE) {
  prepan.pcl.raw.id <- glad.pcl.raw.id
}
```

@@

```{r edgi pcl prepan copy}
if(EDGI == TRUE) {
  prepan.pcl.raw.id <- edgi.pcl.raw.id
}
```

@@

```{r glad phq cols}
if(GLAD == TRUE){
glad.phq.raw <- readRDS(file = paste0(data.raw_path, "/glad/phq_glad.rds"))
dim(glad.phq.raw)
colnames(glad.phq.raw)
}
```

@@

```{r glad phq prepan copy}
if(GLAD == TRUE) {
  prepan.phq.raw.id <- glad.phq.raw.id
}
```

# edgi phq

```{r edgi phq cols}
if(EDGI == TRUE){
  edgi.phq.raw <- readRDS(file = paste0(data.raw_path, "/edgi/phq_edgi.rds"))
  dim(edgi.phq.raw)
  colnames(edgi.phq.raw)
}
```

```{r edgi phq cleaning}
if(EDGI == TRUE){
  edgi.phq.raw.id <- edgi.phq.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.prepandemic = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      phq.little_interest_or_pleasure_in_doing_things_prepan = phq.little_interest_or_pleasure_in_doing_things,
      phq.feeling_down_depressed_or_hopeless_prepan = phq.feeling_down_depressed_or_hopeless,
      phq.staying_asleep_sleeping_trouble_prepan = phq.staying_asleep_sleeping_trouble,
      phq.feeling_tired_or_having_little_energy_prepan = phq.feeling_tired_or_having_little_energy,
      phq.poor_appetite_or_overeating_prepan = phq.poor_appetite_or_overeating,
      phq.feeling_bad_failure_family_prepan = phq.feeling_bad_failure_family,
      phq.trouble_concentrating_reading_newspaper_prepan = phq.trouble_concentrating_reading_newspaper,
      phq.moving_fidgety_noticed_opposite_prepan = phq.moving_fidgety_noticed_opposite,
      phq.dead_hurting_thoughts_prepan = phq.dead_hurting_thoughts
    ) %>%
    add_numeric(., exclude = exclude_cols) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(edgi.phq.raw.id)
  # Inspect colnames
  colnames(edgi.phq.raw.id)
  #Differences
  dim(edgi.phq.raw)[1]-dim(edgi.phq.raw.id)[1]
}
```

```{r edgi phq prepan copy}
if(EDGI == TRUE) {
  prepan.phq.raw.id <- edgi.phq.raw.id
}
```

@@

*Note that scores with missingness will be changed to NA later in the script
```{r phq sum score prepan}
if(GLAD == TRUE | EDGI == TRUE) {
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
phq.items_key_prepan <- c(1,1,1,1,1,1,1,1,1)

@@

# Calulate sum score
phq.scored_items_prepan <- scoreItems(keys = phq.items_key_prepan,
                                        items = prepan.phq.raw.id[phq.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 3)
# Add in column to tibble
prepan.phq.raw.id$phq.sum_score_prepan_raw <- phq.scored_items_prepan$scores

# Look at values
prepan.phq.raw.id %>%
  descr(phq.sum_score_prepan_raw,
        stats = "common"
        )
}
```

@@

prepandemic
```{r phq prepan missingness}
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    na_per_person_phq_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[phq.items_prepan])
      )
  )

prepan.phq.raw.id %>%
  freq(na_per_person_phq_prepan,
       cumul = F)
```

```{r phq prepan missingness 8 items}
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    phq.missing_only_suicide_item_prepan = 
      if_else(
        condition = na_per_person_phq_prepan == 1 & 
          is.na(phq.dead_hurting_thoughts_prepan_numeric), 
        true = 1,
        false = 0,
        missing = NA_real_
      )
  )

prepan.phq.raw.id %>%
  freq(phq.missing_only_suicide_item_prepan,
       cumul = F)
```

all times
```{r phq prepan sum score if missigng items set to NA - all items}
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    phq.sum_score_prepan =
      if_else(condition = na_per_person_phq_prepan > 0,
              true = NA_real_,
              false = phq.sum_score_prepan_raw,
              missing = NA_real_
      )
  )

prepan.phq.raw.id %>%
  descr(phq.sum_score_prepan_raw, stats = "common")

prepan.phq.raw.id %>%
  descr(phq.sum_score_prepan, stats = "common")
```
